### 2013-03-21 inst/examples/pomp-example.Rmd
pomp example
need to combine versions agressively here...

```{r}
library(pomp)
```

@@

```{r}
theta <- c(r = 0.75, K = 10, Q = 3, H = 1, a = 1.3, 
           sigma = 0.1, tau = 0.1, X.0 = 5) 
```
```

@@

```{r}
dat <- simulate(may, params=theta, obs=TRUE, states=TRUE)
library(reshape2)
library(ggplot2)
dat <- melt(dat)[3:5]
names(dat) = c("time", "value", "variable")
ggplot(dat, aes(time, value, lty=variable)) + geom_line()
```

Note that simulate returns an instance of our object, with the data column filled in.  This is not true if we give it arguments such as state or obs = TRUE.

@@

```{r}
measure.dens <- function(y, x, t, params, log, ...) {
    tau <- params["tau"]
    ## state at time t:
    X <- x["X"]
    ## observation at time t:
    Y <- y["Y"]
    ## compute the likelihood of Y|X,tau
    dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)
}
```

@@

```{r}
may <- pomp(may, dmeasure = measure.dens)
```

@@

Run the particle filter

@@

```{r}

```

>>>>

```{r}
library(pomp)
opts_knit$set(upload.fun = socialR::flickr.url)
options(flickr_tags = "nonparametric-bayes")
```

@@

```{r}
theta <- c(r = 0.75, K = 10, Q = 3, H = 1, a = 1.3, 
           sigma = 0.2, tau = 0.1, X.0 = 5) 
```
```

@@

```{r}
dat <- simulate(may, params=theta, obs=TRUE, states=TRUE)
library(reshape2)
library(ggplot2)
dt <- melt(dat)[3:5]
names(dt) = c("time", "value", "variable")
ggplot(dt, aes(time, value, lty=variable)) + geom_line()
```

Note that simulate returns an instance of our object, with the data column filled in.  This is not true if we give it arguments such as state or obs = TRUE as above.  However, when we query for observed and true states in this way, the simulated data is not stored in our object.  We must assign it with `may <- pomp(data = ...` where data is a data frame with columns "time" and "Y".  This is much simpler if we use the default return:

```{r}
may <- simulate(may, params=theta)
```

@@

```{r}
measure.density <- function(y, x, t, params, log, ...) {
    tau <- params["tau"]
    ## state at time t:
    X <- x["X"]
    ## observation at time t:
    Y <- y["Y"]
    ## compute the likelihood of Y|X,tau
    f <- dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)

}
```

@@

```{r}
may <- pomp(may, dmeasure = measure.density)
```

@@

Run the particle filter.  Note that this evaluates the likelihood of our simulated data (stored in the object) at the given (true) parameters.

@@

Here we go, estimate the model by particle filter.  This'll be slow...

```{r}
theta.guess <- theta + abs(rnorm(length(theta)))
estpars <- names(theta) # which pars should be estimated. lets try and get them all
rw.sd <- theta/(50*theta) # a named vector of random walk values. This sets it to 0.02 for all pars
system.time(mf <- 
  mif(may, Nmif = 100, # Number of iterations
      start = theta.guess, # Starting value of parameters
      transform = TRUE, # work in transformed parameter space (e.g. log to avoid negative par values)
      pars = estpars,  # which parameters to estimate (others fixed at starting value)
      rw.sd = rw.sd, # random walk width
      Np = 2000, # Number of particles
      var.factor = 4, # scaling factor for random walk sd
      ic.lag = 10, # fixed lag
      cooling.factor = 0.999, # exponential cooling
      max.fail = 10) # internal tolerance
)
```

```{r}
compare.mif(mf)
```

```{r}
coef(mf)
logLik(mf)
pf_est <- pfilter(may, params = theta.guess, Np = 1000)
logLik(pf_est) # initial loglik
```

These functions apply to replicated mifs above.

```{r}
theta.true <- coef(may)
theta.mif <- apply(sapply(mf,coef),1,mean)
loglik.mif <- replicate(n=10,logLik(pfilter(mf[[1]],params=theta.mif,Np=10000)))
bl <- mean(loglik.mif)
loglik.mif.est <- bl+log(mean(exp(loglik.mif-bl)))
loglik.mif.se <- sd(exp(loglik.mif-bl))/exp(loglik.mif.est-bl)
```
